### 2020-12-28 manuscript/tabs/spt1.tex
make demographics tables and fix manuscript formatting

\begin{table}[ht]
    \centering
    \begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|l|}
    \hline
        Independent Cohort \#1 &  &  &  &  &  &  &  &  &  &  &  \\ \hline
         &  &  &  &  &  &  &  &  &  &  &  \\ \hline
        Source & Dataset name & \# samples &  &  &  & Sex (\% female) &  &  & Age (mean, sd) &  &  \\ \hline
         &  & HLT & NAT & TUM & Total & HLT & NAT* & TUM* & HLT & NAT* & TUM* \\ \hline
        GDC & TCGA-COAD & -- & -- & 135 & 135 & -- & -- & 46.7\% & -- & -- & 69.1, 12.0 \\ \hline
        GDC & TCGA-READ & -- & -- & 60 & 60 & -- & -- & 41.7\% & -- & -- & 67.7, 10.0 \\ \hline
        SRA & MtSinai & -- & 47 & -- & 47 & -- & 61.7\% & -- & -- & 65.4, 13.3 & -- \\ \hline
        SRA & CityOfHope & -- & 2 & 8 & 10 & -- & NR & NR & -- & NR & NR \\ \hline
        SRA & Singapore & -- & 5 & 2 & 7 & -- & NR & NR & -- & NR & NR \\ \hline
        SRA & Utah & 10 & -- & -- & 10 & 40.0\% & -- & -- & 54.8, 7.3 & -- & -- \\ \hline
        SRA & ZhejiangU & -- & 2 & 2 & 4 & -- & 0.0\% & 0.0\% & -- & 58.5, 30.4 & 53.5, 2.1 \\ \hline
        SRA & TexasAM & 2 & -- & -- & 2 & 50.0\% & -- & -- & 38.5, 7.8 & -- & -- \\ \hline
        Total &  & 12 & 56 & 207 & 275 & 41.7\% & 59.2\% & 44.7\% & 52.1, 9.5 & 65.1, 13.8 & 68.5, 11.4 \\ \hline
        * missing data on subset of samples; NR signifies not reported &  &  &  &  &  &  &  &  &  &  &  \\ \hline
    \end{tabular}
    \caption{Supplementary Table 1}
    \label{tab:ind1}
\end{table}