### 2020-12-28 manuscript/tabs/tab_disc.tex

\begin{table}[ht]
    \small
    \centering
    \caption{Phenotypic and demographic characteristics of discovery cohort}
    \label{tab:discovery}
    \begin{tabular}{ll|llll|lll|lll}
        \textbf{Source} & \textbf{Dataset} & \multicolumn{4}{l|}{\textbf{\# samples}} & \multicolumn{3}{l|}{\textbf{Sex (\% female)}} & \multicolumn{3}{l}{\textbf{Age (mean, sd)}} \\
        \cline{3-12}
         & & HLT & NAT & TUM & Total & HLT & NAT* & TUM* & HLT & NAT* & TUM* \\
        \hline
        BarcUVa & BarcUVa & 260 & -- & -- & 260 & 63 & -- & -- & 60, 7 & -- & -- \\
        GDC & TCGA-COAD & -- & 36 & 209 & 245 & -- & 53 & 45 & -- & 72, 13 & 65, 13 \\
        GDC & TCGA-READ & -- & 8 & 81 & 89 & -- & 88 & 46 & -- & 67, 18 & 63, 12 \\
        SRA & GTEx & 202 & -- & -- & 202 & 42 & -- & -- & 49, 13 & -- & -- \\
        SRA & HebeiMU & -- & 4 & 5 & 9 & -- & 75 & 60 & -- & \emph{NR} & \emph{NR} \\
        SRA & KoreaAMC & -- & 8 & 6 & 14 & -- & \emph{NR} & \emph{NR} & -- & \emph{NR} & \emph{NR} \\
        SRA & KoreaPNU & -- & 1 & 2 & 3 & -- & 0 & 100 & -- & 62, -- & 73, 4 \\
        SRA & Mayo & -- & 4 & 8 & 12 & -- & 50 & 38 & -- & 65, 16 & 66, 18 \\
        \hline
        Total &  & 462 & 61 & 311 & 834 & 54\% & 59\% & 46\% & 55, 11 & 70, 14 & 65, 13 \\
        \hline
        \multicolumn{12}{l}{* missing data on subset of samples; \emph{NR} signifies not reported}
    \end{tabular}
\end{table}