### 2020-12-28 manuscript/tabs/tab_ind2.tex

\begin{table}[ht]
    \centering
    \begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|l|}
    \hline
        Independent Cohort \#2 &  &  &  &  &  &  &  &  &  &  &  \\ \hline
         &  &  &  &  &  &  &  &  &  &  &  \\ \hline
        Source & Dataset name & \# samples &  &  &  & Sex (\% female) &  &  & Age (mean, sd) &  &  \\ \hline
         &  & HLT & NAT & TUM & Total & HLT & NAT* & TUM* & HLT & NAT* & TUM* \\ \hline
        GDC & TCGA-COAD & -- & 3 & 3 & 6 & -- & 50.0\% & 50.0\% & -- & 62.2, 13.4 & 62.2, 13.4 \\ \hline
        GDC & TCGA-READ & -- & 1 & 1 & 2 & -- & 0.0\% & 0.0\% & -- & 50.0, -- & 50.0, -- \\ \hline
        SRA & HebeiMU & -- & 1 & 1 & 2 & -- & 0.0\% & 0.0\% & -- & NR & NR \\ \hline
        SRA & KoreaAMC & -- & 4 & 4 & 8 & -- & NR & NR & -- & NR & NR \\ \hline
        SRA & KoreaPNU & -- & 2 & 2 & 4 & -- & 50.0\% & 50.0\% & -- & 74.0, 1.2 & 74.0, 1.2 \\ \hline
        SRA & Mayo & -- & 4 & 4 & 8 & -- & 50.0\% & 50.0\% & -- & 70.3, 9.7 & 70.3, 9.7 \\ \hline
        Total &  & -- & 15 & 15 & 30 & -- & 36.4\% & 36.4\% & -- & 66.6, 11.6 & 66.6, 11.6 \\ \hline
        * missing data on subset of samples; NR signifies not reported &  &  &  &  &  &  &  &  &  &  &  \\ \hline
    \end{tabular}
    \caption{Supplementary Table 2}
    \label{tab:ind2}
\end{table}