### 2020-12-29 manuscript/main.tex
break out manuscript sections and supplement

\documentclass[11pt,letterpaper]{article}

    %% load packages
    \usepackage[margin=1in]{geometry}         %% document margins
    \usepackage[utf8]{inputenc}
    \usepackage{setspace}                     %% to get \doublespacing command
    \usepackage{authblk}                      %% to get \affil command
    \usepackage{bibentry}                     %% bib formatting
    \usepackage{parskip}                      %% space between paragraphs
    \usepackage{fontspec}                     %% for \setmainfont command, will need xelatex to compile Arial
    \usepackage[comma,super,sort&compress]{natbib}  %% https://latex.org/forum/viewtopic.php?t=8399
    \usepackage{graphicx}                     %% for figures
    \usepackage{caption}                      %% for \captionsetup command
    \usepackage{placeins}                     %% for \FloatBarrier command

    %% general document formatting commands
    \pagestyle{plain}                         %% page numbers in middle
    %\linespread{1.67}                        %% baseline linespread = 1.2 (1.2*1.67~2)
    \doublespacing
    \setmainfont{Arial}                       %% https://latex.org/forum/viewtopic.php?t=25998
    \setcitestyle{citesep={,}}
    \graphicspath{{figs/}}                    %% set path to figures relative to this document
    \captionsetup{singlelinecheck=false, justification=justified}

    \title{Potential Novel Immunotarget Revealed in Mega Analysis of Colorectal Transcriptomes}

    \author[1]{Christopher H. Dampier, MD}
    \author[2]{David F. Grabski, MD, PhD}
    \author[1]{Matthew Devall, PhD}
    \author[3]{Virginia Diez-Obrero, BS}
    \author[3]{Ferran Moratalla-Navarro, MS}
    \author[4]{David Rekosh, PhD}
    \author[4]{Marie-Louise Hammarskjold, PhD}
    \author[5]{Sara K. Rasmussen, MD, PhD}
    \author[3]{Victor Moreno, MD, PhD}
    \author[1,*]{Graham Casey, PhD}

    \affil[1]{Center for Public Health Genomics, University of Virginia, Charlottesville, Virginia, USA}
    \affil[2]{Department of Surgery, University of Virginia, Charlottesville, Virginia, USA}
    \affil[3]{Catalan Institute of Oncology, Barcelona, Spain}
    \affil[4]{Department of Microbiology, Immunology and Cancer Biology, University of Virginia, Charlottesville, Virginia, USA}
    \affil[5]{Department of Surgery, Seattle Children's, Seattle, Washington, USA}
    \affil[*]{Correspondence: Graham Casey, PhD, Center for Public Health Genomics, MSB Room 3238, Department of Public Health Sciences, University of Virginia, P.O. Box 800717, Charlottesville, VA 22908-0717, gc8r@virginia.edu}

    %% redefine maketitle function to customize title formatting
        %% https://tex.stackexchange.com/questions/85343/left-align-abstract-title-and-authors
        %% https://en.wikibooks.org/wiki/LaTeX/Fonts#Sizing_text
    \makeatletter
    \renewcommand{\maketitle}{
        \begingroup
            \setlength{\parindent}{0pt}
            \begin{flushleft}
                \LARGE\textbf{\@title}
                \newline
                \newline
                \normalsize\@author
            \end{flushleft}
        \endgroup
    }
    \makeatother

\begin{document}

\maketitle

\newpage
\section*{Synopsis}
\input{secs/synopsis.tex}

\newpage
\section*{Abstract}
\input{secs/abstract.tex}

\newpage
\section*{Introduction}
\input{secs/introduction.tex}

\section*{Methods}
\input{secs/methods.tex}

\section*{Results}
\input{secs/results.tex}

\section*{Discussion}
\input{secs/discussion.tex}

\newpage
\section*{Acknowledgements}

People we want to thank. Grant support.

\newpage
\bibliography{main}
\bibliographystyle{myvancouver}

\newpage
\section*{Figure Legends}
\input{secs/legends.tex}

\newpage
\section*{Tables}
\input{tabs/tab1.tex}
\input{tabs/tab2.tex}

\FloatBarrier
\newpage
\section*{Figures}

\begin{figure}[ht]
  \makebox[\textwidth][c]
  {
    \includegraphics[width=0.9\textwidth]{fig1}
  }
  \caption{Summary of unannotated human endogenous retroviral sequences}
  \label{fig:pre_sum}
\end{figure}

\begin{figure}[ht]
  \makebox[\textwidth][c]
  {
    \includegraphics[width=0.9\textwidth]{fig2}
  }
  \caption{Differential expression results for unannotated HERV genes}
  \label{fig:dge_res}
\end{figure}

\begin{figure}[ht]
  \makebox[\textwidth][c]
  {
    \includegraphics[width=0.9\textwidth]{fig3}
  }
  \caption{Global estimates of HERV expression across tissue phenotypes}
  \label{fig:heat}
\end{figure}

\end{document}

problems with AMA style file
https://tex.stackexchange.com/questions/55020/updated-ama-bst-file
i modified vancouver.bst to get proper author lists