### 2020-12-29 manuscript/secs/legends.tex

\textbf{Figure 1.} Overview of human endogenous retroviral element discovery in colorectal cancer.
A, Barplot demonstrating
B, Barplot demonstrating
C, Histogram demonstrating
D, Histogram demonstrating

\textbf{Figure 2.} Unannotated proviral elements demonstrate tissue phenotype specificity in colorectal transcriptomes.
A, Scatterplot (i.e. volcano plot) demonstrating
B, Scatterplots (i.e. volcano plot) demonstrating
C, Boxplots demonstrating

\textbf{Figure 3.} Global tendency of healthy colorectal transcriptomes to harbor relatively higher expression of unannotated proviral elements.
A, Heatmap demonstrating
B, Heatmap demonstrating