### 2020-12-31 manuscript/tabs/tab_ind1.tex

\begin{table}[ht]
    \small
    \centering
    \caption{Phenotypic and demographic characteristics of independent cohort \#1}
    \label{tab:ind1}
    \begin{tabular}{ll|llll|lll|lll}
        \textbf{Source} & \textbf{Dataset} & \multicolumn{4}{l|}{\textbf{\# samples}} & \multicolumn{3}{l|}{\textbf{Sex (\% female)}} & \multicolumn{3}{l}{\textbf{Age (mean, sd)}} \\
        \cline{3-12}
         & & HLT & NAT & TUM & Total & HLT & NAT* & TUM* & HLT & NAT* & TUM* \\
        \hline
        GDC & TCGA-COAD & -- & -- & 135 & 135 & -- & -- & 47 & -- & -- & 69, 12 \\
        GDC & TCGA-READ & -- & -- & 60 & 60 & -- & -- & 42 & -- & -- & 68, 10 \\
        SRA & MtSinai & -- & 47 & -- & 47 & -- & 62 & -- & -- & 65, 13 & -- \\
        SRA & CityOfHope & -- & 2 & 8 & 10 & -- & \emph{NR} & \emph{NR} & -- & \emph{NR} & \emph{NR} \\
        SRA & Singapore & -- & 5 & 2 & 7 & -- & \emph{NR} & \emph{NR} & -- & \emph{NR} & \emph{NR} \\
        SRA & Utah & 10 & -- & -- & 10 & 40 & -- & -- & 55, 7 & -- & -- \\
        SRA & ZhejiangU & -- & 2 & 2 & 4 & -- & 0 & 0 & -- & 59, 30 & 54, 2 \\
        SRA & TexasAM & 2 & -- & -- & 2 & 50 & -- & -- & 39, 8 & -- & -- \\
        \hline
        Total &  & 12 & 56 & 207 & 275 & 42\% & 59\% & 45\% & 52, 10 & 65, 14 & 69, 11 \\
        \hline
        \multicolumn{12}{l}{* missing data on subset of samples; \emph{NR} signifies not reported} \\
    \end{tabular}
\end{table}