### 2020-12-31 manuscript/tabs/tab_ind2.tex

\begin{table}[ht]
    \small
    \centering
    \caption{Phenotypic and demographic characteristics of independent cohort \#2}
    \label{tab:ind2}
    \begin{tabular}{ll|llll|lll|lll}
        \textbf{Source} & \textbf{Dataset} & \multicolumn{4}{l|}{\textbf{\# samples}} & \multicolumn{3}{l|}{\textbf{Sex (\% female)}} & \multicolumn{3}{l}{\textbf{Age (mean, sd)}} \\
        \cline{3-12}
         & & HLT & NAT & TUM & Total & HLT & NAT* & TUM* & HLT & NAT* & TUM* \\
        \hline
        GDC & TCGA-COAD & -- & 3 & 3 & 6 & -- & 50 & 50 & -- & 62, 13 & 62, 13 \\
        GDC & TCGA-READ & -- & 1 & 1 & 2 & -- & 0 & 0 & -- & 50, -- & 50, -- \\
        SRA & HebeiMU & -- & 1 & 1 & 2 & -- & 0 & 0 & -- & \emph{NR} & \emph{NR} \\
        SRA & KoreaAMC & -- & 4 & 4 & 8 & -- & \emph{NR} & \emph{NR} & -- & \emph{NR} & \emph{NR} \\
        SRA & KoreaPNU & -- & 2 & 2 & 4 & -- & 50 & 50 & -- & 74, 1 & 74, 1 \\
        SRA & Mayo & -- & 4 & 4 & 8 & -- & 50 & 50 & -- & 70, 10 & 70, 10 \\
        \hline
        Total &  & -- & 15 & 15 & 30 & -- & 36\% & 36\% & -- & 67, 12 & 67, 12 \\
        \hline
        \multicolumn{12}{l}{* missing data on subset of samples; \emph{NR} signifies not reported} \\
    \end{tabular}
\end{table}