### 2020-12-31 manuscript/secs/acknowledge.tex

The results shown here are in part based on data generated by The Cancer Genome Atlas (TCGA) Research Network: https://www.cancer.gov/tcga.
The results are also based on data generated by the Genotype-Tissue Expression (GTEx) Project.
GTEx was supported by the Common Fund of the Office of the Director of the National Institutes of Health and by NCI, NHGRI, NHLBI, NIDA, NIMH, and NINDS.
The data used for the analyses described in this manuscript were obtained from dbGaP accession number phs000424.v7.p2 on2020-07-15.
This research was supported by the Genomics Core Facility of the CWRU School of Medicine's Genetics and Genome Sciences Department.
Funding for this work was provided by the following NIH Grants: R01 NIH/NCI CA143237 and CA201407 (G.C.), T32 5T32CA163177-07 (Craig Slingluff, MD).
The authors are grateful to Dr Craig Slingluff from the Department of Surgery at the University of Virginia, whose T32 training grant supported CHD and DFG.