### 2021-01-24 manuscript/main.tex
first set of revisions to manuscript, excluding MAD

\documentclass[11pt,letterpaper]{article}  %% 3500 words for ASO research article

    %% load packages
    \usepackage[margin=1in]{geometry}         %% document margins
    \usepackage[utf8]{inputenc}
    \usepackage{setspace}                     %% to get \doublespacing command
    \usepackage{authblk}                      %% to get \affil command
    \usepackage{bibentry}                     %% bib formatting
    \usepackage{parskip}                      %% space between paragraphs
    \usepackage{fontspec}                     %% for \setmainfont command, will need xelatex to compile Arial
    \usepackage[comma,super,sort&compress]{natbib}  %% https://latex.org/forum/viewtopic.php?t=8399
    \usepackage{graphicx}                     %% for figures
    \usepackage{caption}                      %% for \captionsetup command
    \usepackage{placeins}                     %% for \FloatBarrier command

    %% general document formatting commands
    \pagestyle{plain}                         %% page numbers in middle
    %\linespread{1.67}                        %% baseline linespread = 1.2 (1.2*1.67~2)
    \doublespacing
    \setmainfont{Arial}                       %% https://latex.org/forum/viewtopic.php?t=25998
    \setcitestyle{citesep={,}}
    \graphicspath{{figs/}}                    %% set path to figures relative to this document
    \captionsetup{singlelinecheck=false, justification=justified}

    %\title{Potential Novel Immunotarget Revealed in Mega Analysis of Colorectal Transcriptomes}
    % \title{Human Endogenous Retroviral Transcripts Show Potential as Novel, Tumor Specific Biomarkers for Colorectal Cancer}
    \input{secs/title.tex}

    %% redefine maketitle function to customize title formatting
        %% https://tex.stackexchange.com/questions/85343/left-align-abstract-title-and-authors
        %% https://en.wikibooks.org/wiki/LaTeX/Fonts#Sizing_text
    \makeatletter
    \renewcommand{\maketitle}{
        \begingroup
            \setlength{\parindent}{0pt}
            \begin{flushleft}
                \LARGE\textbf{\@title}
                \newline
                \newline
                \small\@author
            \end{flushleft}
        \endgroup
    }
    \makeatother

\begin{document}

\maketitle
\small\textbf{Running Header}: HERV Transcripts in Colorectal Tumors \\
\small\textbf{Disclosure Statement}: No author has any commercial interest in the subject of study.

\newpage
\section*{Synopsis}
\input{secs/synopsis.tex}  %% 1-3 sentences, 40 words max

\newpage
\section*{Abstract}
\input{secs/abstract.tex}

\newpage
\section*{Introduction}
\input{secs/introduction.tex}

\section*{Methods}
\input{secs/methods.tex}

\section*{Results}
\input{secs/results.tex}

\section*{Discussion}
\input{secs/discussion.tex}

\newpage
\section*{Acknowledgements}
\input{secs/acknowledge.tex}

\newpage
\bibliography{main}
\bibliographystyle{myvancouver}

\newpage
\section*{Figure Legends}
\input{secs/legends.tex}

\newpage
\section*{Tables}
\input{tabs/tab1.tex}
\input{tabs/tab2.tex}

\FloatBarrier
\newpage
\section*{Figures}

\begin{figure}[ht]
  \makebox[\textwidth][c]
  {
    \includegraphics[width=0.9\textwidth]{fig1}
  }
  \caption{Summary of unannotated human endogenous retroviral sequences found in colorectal transcriptomes}
  \label{fig:pre_sum}
\end{figure}

\begin{figure}[ht]
  \makebox[\textwidth][c]
  {
    \includegraphics[width=0.9\textwidth]{fig2}
  }
  \caption{Differential expression of unannotated human endogenous retroviral loci in colorectal transcriptomes}
  \label{fig:dge_res}
\end{figure}

\begin{figure}[ht]
  \makebox[\textwidth][c]
  {
    \includegraphics[width=0.9\textwidth]{fig3}
  }
  \caption{Global estimates of unannotated human endogenous retroviral expression across tissue phenotypes.}
  \label{fig:heat}
\end{figure}

\end{document}

%% citation issues
- problems with AMA style file
- https://tex.stackexchange.com/questions/55020/updated-ama-bst-file
- i modified vancouver.bst to get proper author lists > myvancouver.bst
- pandoc conversion to docx requires csl
- wget -O ama.csl https://github.com/citation-style-language/styles/blob/master/american-medical-association.csl?raw=true

%% word format issues
- see https://pandoc.org/MANUAL.html, --reference-doc= for DOCX
- To produce a custom reference.docx, first get a copy of the default reference.docx:
pandoc --print-default-data-file reference.docx > custom-reference.docx
Then open custom-reference.docx in Word, modify the styles as you wish, and save the file.
Modified: Normal, Abstract, Author, Title, Subtitle, Heading 1, Heading 2, Bibliography, Body Text, First Paragraph, Verbatim Char