### 2021-01-24 manuscript/secs/acknowledge.tex

The results shown here are in part based on data generated by The Cancer Genome Atlas (TCGA) Research Network: https://www.cancer.gov/tcga.
The results are also based on data generated by the Genotype-Tissue Expression (GTEx) Project.
GTEx was supported by the Common Fund of the Office of the Director of the National Institutes of Health and by NCI, NHGRI, NHLBI, NIDA, NIMH, and NINDS.
The data used for the analyses described in this manuscript were obtained from dbGaP accession number phs000424.v7.p2 on 2020-07-15.
This research was supported by the Genomics Core Facility of the CWRU School of Medicine's Genetics and Genome Sciences Department.
Funding for this work was provided by the following grants: NIH/NCI R01 CA143237 (GC), NIH/NCI R01 CA201407 (GC), NIH/NCI R01 CA206275 (DR and MLH), NIH/NCI T32 5T32CA163177-07 (Craig Slingluff, MD), Agency for Management of University and Research Grants of the Catalan Government grant 2017SGR723 (VM), Instituto de Salud Carlos III (co-funded by European Regional Development Fund "A way to build Europe" grants PI14-00613 and PI17-00092), and Spanish Association Against Cancer Scientific Foundation grant GCTRA18022MORE (VM).
The authors are grateful to CERCA Programme, Generalitat de Catalunya for institutional support.
The authors are also grateful to Dr Craig Slingluff from the Department of Surgery at the University of Virginia, whose T32 training grant provided salary support for CHD and DFG.
Partial salary support for DR and MLH was provided by the Charles H. Ross Jr and Myles H. Thaler Professorship endowments at the University of Virginia.