### 2021-01-24 manuscript/secs/title.tex

\title{Human Endogenous Retroviral Transcripts Show Potential as Novel, Tumor Specific Biomarkers for Colorectal Cancer}

\author[1,2]{Christopher H. Dampier, MD}
\author[3,4]{David F. Grabski, MD}
\author[1,2]{Matthew Devall, PhD}
\author[5]{Virginia Diez-Obrero, BS}
\author[5]{Ferran Moratalla-Navarro, MS}
\author[4,6]{David Rekosh, PhD}
\author[4,6]{Marie-Louise Hammarskjold, MD, PhD}
\author[7]{Sara K. Rasmussen, MD, PhD}
\author[5,8,9,10]{Victor Moreno, MD, PhD}
\author[1,2,*]{Graham Casey, PhD}

\affil[1]{Center for Public Health Genomics, University of Virginia, Charlottesville, Virginia, USA}
\affil[2]{Department of Public Health Sciences, University of Virginia, Charlottesville, Virginia, USA}
\affil[3]{Department of Surgery, University of Virginia, Charlottesville, Virginia, USA}
\affil[4]{Thaler Center for HIV and Retrovirus Research, University of Virginia, Charlottesville, Virginia, USA}
\affil[5]{Unit of Biomarkers and Susceptibility, Oncology Data Analytics Program, Catalan Institute of Oncology, Hospitalet de Llobregat, Barcelona, Spain}
\affil[6]{Department of Microbiology, Immunology and Cancer Biology, University of Virginia, Charlottesville, Virginia, USA}
\affil[7]{Department of Surgery, Seattle Children's, Seattle, Washington, USA}
\affil[8]{Colorectal Cancer Group, ONCOBELL Program, Bellvitge Biomedical Research Institute, Hospitalet de Llobregat, Barcelona, Spain}
\affil[9]{Consortium for Biomedical Research in Epidemiology and Public Health, Spain}
\affil[10]{Department of Clinical Sciences, Faculty of Medicine, University of Barcelona, Barcelona, Spain}
\affil[*]{Correspondence: Graham Casey, PhD, Center for Public Health Genomics, MSB Room 3238, Department of Public Health Sciences, University of Virginia, P.O. Box 800717, Charlottesville, VA 22908-0717, gc8r@virginia.edu}