### 2021-01-24 manuscript/tabs/spt3.tex
add supp table with HERV coordinates

% latex table generated in R 4.0.0 by xtable 1.8-4 package
% Sun Jan 24 12:10:28 2021
\begin{table}[ht]
\centering
\caption{Tumor specific HERV proviruses validated in independent cohort} 
\label{tab:valsupp}
\begingroup\footnotesize
\begin{tabular}{lllll}
 {\textbf{HERV ID}} & {\textbf{HERV}} & {\textbf{CHROM}} & {\textbf{START}} & {\textbf{STOP}} \\ 
  \hline
LTR/ERV1$|$HERV1\_I-int\~{}HERV1\_LTRc$|$3$|$34679071$|$34687025 & HERV1\_I 3p22.3 & 3 & 34679071 & 34687025 \\ 
  LTR/ERV1$|$HERVIP10F-int\~{}LTR10A$|$9$|$40746806$|$40752615 & HERVIP10F 9p11.2 & 9 & 40746806 & 40752615 \\ 
  LTR/ERV1$|$HERVIP10F-int\~{}LTR10A$|$9$|$64870413$|$64876219 & HERVIP10F 9q13 & 9 & 64870413 & 64876219 \\ 
  LTR/ERVK$|$HERVK9-int\~{}MER9a2$|$13$|$19626322$|$19631229 & HERVK9 13q12.11 & 13 & 19626322 & 19631229 \\ 
  LTR/ERV1$|$HERVH-int\~{}LTR7Y$|$13$|$109265521$|$109271116 & HERVH 13q33.3 & 13 & 109265521 & 109271116 \\ 
  LTR/ERV1$|$HERVH-int\~{}LTR7B$|$20$|$19752049$|$19756776 & HERVH 20p11.23 & 20 & 19752049 & 19756776 \\ 
  LTR/ERV1$|$HERVH-int\~{}LTR7Y$|$X$|$4540474$|$4546320 & HERVH Xp22.32 & X & 4540474 & 4546320 \\ 
   \hline
\end{tabular}
\endgroup
\end{table}