### 2021-01-24 manuscript/supp.tex
add supp fig legends

\documentclass[11pt,letterpaper]{article}

    %% load packages
    \usepackage[margin=1in]{geometry}         %% document margins
    \usepackage[utf8]{inputenc}
    \usepackage{setspace}                     %% to get \doublespacing command
    \usepackage{authblk}                      %% to get \affil command
    \usepackage{bibentry}                     %% bib formatting
    \usepackage{parskip}                      %% space between paragraphs
    \usepackage{fontspec}                     %% for \setmainfont command, will need xelatex to compile Arial
    \usepackage[comma,super,sort&compress]{natbib}  %% https://latex.org/forum/viewtopic.php?t=8399
    \usepackage{graphicx}                     %% for figures
    \usepackage{caption}                      %% for \captionsetup command
    \usepackage{placeins}                     %% for \FloatBarrier command

    %% general document formatting commands
    \pagestyle{plain}                         %% page numbers in middle
    %\linespread{1.67}                        %% baseline linespread = 1.2 (1.2*1.67~2)
    \doublespacing
    \setmainfont{Arial}                       %% https://latex.org/forum/viewtopic.php?t=25998
    \setcitestyle{citesep={,}}
    \graphicspath{{figs/}}                    %% set path to figures relative to this document
    \captionsetup{singlelinecheck=false, justification=justified}

    \input{sups/title.tex}

    %% redefine maketitle function to customize title formatting
        %% https://tex.stackexchange.com/questions/85343/left-align-abstract-title-and-authors
        %% https://en.wikibooks.org/wiki/LaTeX/Fonts#Sizing_text
    \makeatletter
    \renewcommand{\maketitle}{
        \begingroup
            \setlength{\parindent}{0pt}
            \begin{flushleft}
                \LARGE\textbf{\@title}
                \newline
                \newline
                \small\@author
            \end{flushleft}
        \endgroup
    }
    \makeatother

\begin{document}

\maketitle

\newpage
\section*{Supplementary Methods}
\input{sups/supmethods.tex}

\newpage
\bibliography{main}
\bibliographystyle{myvancouver}

\newpage
\section*{Supplementary Figure Legends}
\input{sups/suplegends.tex}

\newpage
\section*{Supplementary Tables}
\input{tabs/spt1.tex}
\input{tabs/spt2.tex}
\input{tabs/spt3.tex}

\FloatBarrier
\newpage
\section*{Supplementary Figures}

\begin{figure}[ht]
  \makebox[\textwidth][c]
  {
    \includegraphics[width=0.8\textwidth]{spf1}
  }
  \caption{Correlations of effect sizes across studies and cohorts}
  \label{fig:corrs}
\end{figure}

\begin{figure}[ht]
  \makebox[\textwidth][c]
  {
    \includegraphics[width=1.2\textwidth]{spf2}
  }
  \caption{Receiver operating characteristic curves in discovery cohort}
  \label{fig:rocA}
\end{figure}

\begin{figure}[ht]
  \makebox[\textwidth][c]
  {
    \includegraphics[width=0.8\textwidth]{spf3}
  }
  \caption{Relationship between subject age and HERV expression in discovery cohort}
  \label{fig:age}
\end{figure}

\begin{figure}[ht]
  \makebox[\textwidth][c]
  {
    \includegraphics[width=1.2\textwidth]{spf4}
  }
  \caption{Relationship between tumor MSI status or stage and HERV expression in discovery cohort}
  \label{fig:msistage}
\end{figure}

\begin{figure}[ht]
  \makebox[\textwidth][c]
  {
    \includegraphics[width=0.6\textwidth]{spf5}
  }
  \caption{Relationship between cytotoxic activity and HERV expression in discovery cohort}
  \label{fig:cd8}
\end{figure}

\begin{figure}[ht]
  \makebox[\textwidth][c]
  {
    \includegraphics[width=1.2\textwidth]{spf6}
  }
  \caption{Receiver operating characteristic curves in independent cohort}
  \label{fig:rocC}
\end{figure}

\begin{figure}[ht]
  \makebox[\textwidth][c]
  {
    \includegraphics[width=0.9\textwidth]{spf7}
  }
  \caption{Boxplots re-demonstrating healthy specific over-expression using a novel HERV-K transcriptome}
  \label{fig:grabski_method}
\end{figure}

\end{document}